FLM for growth as function of SPEI

### 2020-11-10 analysis/paper/vital rate models.Rmd
reorganized files to make R CMD check happier
added Xavier gridded data SPEI.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(mgcv)
library(dlnm)
library(bbmle)
library(lubridate)
library(glue)
library(patchwork)

@@

source(here("R", "crossbasis.R"))
source(here("R", "utils.R"))
```

@@

- ~~Remove seedlings from dataset, for now~~
- ~~Separate models for CF and fragment, for now~~
- Add model for probability of reproduction and other vital rates
- Model seedlings separately?
- Add random effects, interactions, other complexities
- Explore FGAM as alternative to DLNM (seems like they are doing the same thing, but maybe FGAM allows modeling of interactions?)

- Start building IPM for subset of Heliconia data (in separate file)

@@

```{r data, echo=TRUE}
spei <- read_csv(here("analysis", "data", "derived_data", "spei_lags.csv"))
demog <- read_rds(here("analysis", "data", "derived_data", "ha_survey.rds"))
```

## Data Dictionary

`spei`: lagged SPEI for each site cluster

- `site` (character): site cluster, one of "colosso_clust", "dimona", "km_clust", or "porto_alegre"
- `census_year` (numeric): the census year to join with.  
- `spei_1` (numeric): SPEI value for January of the census year
- `spei_2` (numeric): SPEI value for December of the previous year
- `spei_3` (numeric): SPEI value for November of the previous year
- `spei_4`-- `spei_24`: ... and so on out to January of 2 years prior.

`demog`: demographic surveys of *Heliconia acuminata*

- `plot` (character): either `bdffp_reserve_no` or continuous forest indicated by "-CF"
- `habitat` (character): either 1-ha, 10-ha fragments or continuous forest "CF"
- `ranch` (character): DIM = Dimona, EST = Estaciao, PAL = Porto Alegre
- `bdffp_reserve_no` (character): Official bdffp reserve number, when there is one.
- `ha_id_number` (character): Unique plant ID
- `tag_number` (character): Plant ID within plot:year.  May have changed over lifetime of plant.
- `row` (character): Location in plot
- `column` (numeric): Location in plot
- `year` (integer): Survey year
- `ht` (numeric): plant height
- `shts` (integer): number of shoots
- `infl` (integer): number of inflorescences
- `code_notes` (character): Important ones: "sdlg (1)" = seedling, "dead (2)" = dead
- `code2` (character): "initial.yr.tag" = first year that plant is in the study.  May not necessarily be a seedling.
- `surv` (integer): did the plant survive to the next year? 1 = survived, 0 = died, NA = last year of data, fate unknown.

# Join data sets

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(mgcv)
library(dlnm)
library(bbmle)
library(lubridate)
library(glue)
library(patchwork)
library(sf)

@@

# library(HeliconiaDemography) #load functions in this compendium
load_all()
```

@@

- 
- ~~Remove seedlings from dataset, for now~~
- ~~Separate models for CF and fragment, for now~~
- Add model for probability of reproduction and other vital rates
- Model seedlings separately?
- Add random effects, interactions, other complexities
- Match SPEI data to demographic plots programtically by lat lon data

@@

```{r data, echo=TRUE}
spei <- read_csv(here("analysis", "data", "derived_data", "spei_lags.csv"))
xa <- read_rds(here("analysis", "data", "derived_data", "xa_spei_lags.rds"))
demog <- read_rds(here("analysis", "data", "derived_data", "ha_survey.rds"))
plot_coords <- read_csv(here("analysis", "data", "raw_data", "plot_coords.csv"))
```
# Join data sets

## Gridded data
Get only rows for february (the month of the census), and year for joining

```{r}
xa2 <- 
  xa %>%
  filter(month(yearmonth) == 2) %>% 
  mutate(year = year(yearmonth), .after =yearmonth) %>% 
  separate(latlon, into = c("lat", "lon"), sep = "_")

```

Now, for the demography data, get year and plot coords

```{r}
# demog$plot %>% unique() %>% sort()
# plot_coords$plot %>% unique() %>% sort()
demog <-
  right_join(plot_coords %>% select(plot, lon, lat),
             demog,
             by = "plot") %>% 
  select(-x_09, -y_09) #remove to avoid confusion since i'm not using here.
```

Now join by year and whether plot is in grid cell or not.  Probably need to construct `sf` object for grid cells (knowing that the lat lon are the centers and the resolution is 0.25º) and then use some `sf_` function to check if a point is inside of a grid cell.  See: https://stackoverflow.com/questions/59766153/left-join-based-on-closest-lat-lon-in-r

## Imputed data:
